---
title: "An Overview of the Tidyverse"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
  github_document: default
---


This document is based on the Michael Levy's presentation at the Davis R-Users' Group. 

- The youtube video of the presentation is available [here.](https://www.youtube.com/watch?v=_rPhSAVhs1A) 
- The original github repository of the presentation is [here.](https://github.com/michaellevy/tidyverse_talk)

## What is the tidyverse?

A a suite of R tools that follow a tidy philosophy:

### Tidy Philosophy

Put data in data frames  

- Each variable gets a column
- Each observation gets a row
- Each unit of analysis gets a data frame

### Tidy APIs

Functions should be consistent and easily (human) readable

- Take one step at a time
- Connect simple steps with the pipe
- Referential transparency


### Okay but really, what is it? 

Suite of ~20 packages that provide consistent, user-friendly, smart-default tools to do most of what most people do in R.

- Core packages: ggplot2, dplyr, tidyr, readr, purrr, tibble
- Data import: DBI, haven, httr, jsonlite, readxl, rvest, xml2
- Specialized data manipulation: hms, stringr, lubridate, forcats
- Modeling: modelr, broom

`install.packages(tidyverse)` installs all of the above packages.

`library(tidyverse)` attaches only the core packages.

```{r Attach core packages, echo=TRUE}
library(tidyverse)
```

## tibble

A modern reimagining of a data frame.

```{r class tbl}
tdf <- tibble(x = 1:1e4, y = rnorm(1e4))
class(tdf)
```

Tibbles print politely. 

```{r print tdl}
tdf
```


- Can customize print methods with `print(tdf, n = rows, width = cols)`

- Set default with `options(tibble.print_max = rows, tibble.width = cols)`

Tibbles have some convenient and consistent defaults that are different from base R data.frames. 

- In tibbles strings are NOT automatically reconized as factors

Also note that tidyverse import functions, such as `readr::read_csv`, default to tibbles and that *this can break existing code*.

## The pipe `%>%` : Functional composition

Sends the output of the LHS function to the first argument of the RHS function.

```{r pipe, echo=TRUE}
sum(1:8) %>% 
  sqrt()
```

Note that keyboard shortcut for the pipe is `cmd + shift + M`

## dplyr

A package for data manipulation
















